\bibliographystyle{IEEEtran}

\begin{thebibliography}{00}

\bibitem{3gpp-23501}
3GPP, ``System architecture for the 5G System (5GS),'' 3GPP TS 23.501, 2023.

\bibitem{3gpp-23502}
3GPP, ``Procedures for the 5G System (5GS),'' 3GPP TS 23.502, 2023.

\bibitem{3gpp-38413}
3GPP, ``NG-RAN; NG Application Protocol (NGAP),'' 3GPP TS 38.413, 2023.

\bibitem{3gpp-24501}
3GPP, ``Non-Access-Stratum (NAS) protocol for 5G System (5GS),'' 3GPP TS 24.501, 2023.

\bibitem{open5gs}
S. Lee, ``Open5GS: Open Source Project of 5GC and EPC,'' 2023. [Online]. Available: https://open5gs.org/

\bibitem{free5gc}
free5GC, ``free5GC: The free 5G Core Network,'' 2023. [Online]. Available: https://free5gc.org/

\bibitem{ueransim}
A. Heghedus, ``UERANSIM: Open Source 5G UE and RAN (gNB) Implementation,'' 2023. [Online]. Available: https://github.com/aligungr/UERANSIM

\bibitem{free5gc-ngap}
free5GC, ``NGAP: NG Application Protocol Library,'' 2023. [Online]. Available: https://github.com/free5gc/ngap

\bibitem{snf-2020}
A. Singhvi, A. Balasubramanian, K. Houck, M. D. Shaikh, S. Venkataraman, and A. Akella, ``Serverless Network Functions,'' in \textit{Proc. ACM SoCC}, 2020.

\bibitem{5gmedia-2018}
D. Breitgand, ``Towards Serverless NFV for 5G Media Applications,'' in \textit{Proc. ACM SYSTOR}, 2018.

\bibitem{qfaas-2022}
K. Hou, S. Lin, Y. Chen, and V. Yegneswaran, ``QFaaS: Accelerating and Securing Serverless Cloud Networks with QUIC,'' in \textit{Proc. ACM SoCC}, 2022.

\bibitem{directfaas-2024}
Q. Zeng, K. Hou, X. Leng, and Y. Chen, ``DirectFaaS: A Clean-Slate Network Architecture for Efficient Serverless Chain Communications,'' in \textit{Proc. ACM WWW}, 2024.

\bibitem{cloud5gc-2024}
K. Watanabe, K. Shima, K. Horiba, and Y. Sekiya, ``Cloud5GC: Design and Implementation of Scalable and Stateless Mobile Core System on Public Cloud,'' in \textit{Proc. IEEE ICOIN}, 2024.

\bibitem{pp5gs-2023}
S. Goshi, F. Stahl, et al., ``PP5GS---An Efficient Procedure-Based and Stateless Architecture for Next-Generation Core Networks,'' \textit{IEEE Trans. Netw. Service Manag.}, 2023.

\bibitem{proc-decomp-2022}
A. Barakabitze et al., ``Procedure-based Functional Decomposition for 5G Core Network Functions,'' Universit\"{a}t T\"{u}bingen, 2022.

\bibitem{oran-bench-2024}
M. S. Ali et al., ``Control Plane Performance Benchmarking and Feature Analysis of Popular Open-Source 5G Core Networks,'' \textit{IEEE Access}, vol. 12, 2024.

\bibitem{5gc-bench-2023}
A. Mamushiane et al., ``Evaluating the Performance of Open Source Software Implementations of the 5G Network Core,'' in \textit{Proc. IEEE NOMS}, 2023.

\bibitem{ionos-pricing}
IONOS SE, ``Compute Engine Pricing,'' 2025. [Online]. Available: https://cloud.ionos.com/prices

\bibitem{aws-lambda-pricing}
Amazon Web Services, ``AWS Lambda Pricing,'' 2025. [Online]. Available: https://aws.amazon.com/lambda/pricing/

\end{thebibliography}

\balance

\end{document}